\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-03-12 manuscript/manuscript.tex
sp100 and sp500 tables

The Kolmogorov-Smirnov (KS) test is a useful goodness-of-fit statistic. 
Let $X_1, ..., X_n$ be a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ be that $X_i$'s follow the
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. The distribution function can be
accurately evaluated in modern statistical software
\citep{marsaglia2003evaluating}. The KS test can be applied to a variety of
fields. \jy{cite a few  references in various field to show its wide
  applications.}
 It has
been used to analyze the random distribution of cosmic microwave background 
radiation \citep{naess2012application}.

Despite its widespread use, the Kolmogorov-Smirnov (KS) test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
Another commonly encountered situation is when the data are serially
dependent. \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods. Alternatively, 
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$, Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
distributions.
The autocorrelation $\phi$ of the series was set to
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$. These
are the autocorrelations of the normal series, but after the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.607, 
diverges, -0.287, 0, 0.339, 0.668, 0.911\}$.
\mc{my explanation needs work}

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. The series were generated by marginally 
transforming the
marginal normal AR(1) series $X_t$ by
\begin{equation*}
W_t = F^{-1}[\Phi(X_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $\Gamma(8, 1)$.
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is normal, and when $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is gamma. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

\jy{Consider presenting this in a graph. Try explaining the pattern.
  Keep the notations consistent: $n$ and n are different.}

@@

\subsection{Precipitation at Airports}
\label{sec:precipitation}
Annual maximums of hourly precipitation in inches time series data were 
retrieved 
from the Chicago Midway International Airport (MDW), New York LaGuardia 
International 
Airport (LGA),
and Los Angeles International Airport (LAX) stations from the 
National Oceanic and Atmospheric Administration (NOAA). These stations were selected 
from places of importance in the three largest American cities. LGA and
MDW were chosen over John F. Kennedy and O'Hare as more data was available
from these stations. The data for MDW is from
1948 to 2014, the data for LGA is from 1948 to 2013, and the
data for LAX is also from 1948 to 2013. We can apply our method (as well
as the non-parametric bootstrap KS Test \citep{babu2004goodness} and
the parametric bootstrap KS Test) to
the precipitation time series from each of these locations to test if they
follow the
Generalized Extreme Value (GEV) distribution using the \textsl{evd} 
package \citep{evd}. Precipitation maximum series 
typically do follow the GEV distribution, so we expect to fail to reject the
null hypothesis.

\jy{Is the parametric method by Zeimberkakis?}

@@

>>>>

The Kolmogorov-Smirnov (KS) test is a useful goodness-of-fit statistic. 
Let $X_1, ..., X_n$ be a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ be that $X_i$'s follow the
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. The distribution function can be
accurately evaluated in modern statistical software
\citep{marsaglia2003evaluating}. The KS test can be applied to a variety of
fields. \jy{cite a few  references in various field to show its wide
  applications.}
 It has
been used to analyze the random distribution of cosmic microwave background 
radiation \citep{naess2012application}. \citep{aslam2020introducing} 
applied the method to test if the count rate per second of
radioactive data followed the Poisson distribution. 
Researchers have proposed the
method as a technique for gear condition monitoring \citep{andrade2001gear}. It
can also be used to compare images of breast
cancer tumors \citep{demidenko2004kolmogorov}.

Despite its widespread use, the Kolmogorov-Smirnov (KS) test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from samples generated from the fitted hypothesized distribution.
Another commonly encountered situation is when the data are serially
dependent. \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods. Alternatively, 
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$, Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
distributions.
We generated time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming the $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(X_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the unit exponential 
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.607, 
diverges, -0.287, 0, 0.339, 0.668, 0.911\}$.
\mc{my explanation needs work}

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is normal, and when $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is gamma. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

\jy{Consider presenting this in a graph. Try explaining the pattern.
  Keep the notations consistent: $n$ and n are different.}
\mc{addressed}

@@

\subsection{Precipitation at Airports}
\label{sec:precipitation}
Annual maximums of hourly precipitation in inches time series data were 
retrieved 
from the Chicago Midway International Airport (MDW), New York LaGuardia 
International 
Airport (LGA),
and Los Angeles International Airport (LAX) stations from the 
National Oceanic and Atmospheric Administration (NOAA). These stations were selected 
from places of importance in the three largest American cities. LGA and
MDW were chosen over John F. Kennedy and O'Hare as more data was available
from these stations. The data for MDW is from
1948 to 2014, the data for LGA is from 1948 to 2013, and the
data for LAX is also from 1948 to 2013. We can apply our method (as well
as the non-parametric bootstrap KS Test \citep{babu2004goodness} and
the parametric bootstrap KS Test \citep{zeimbekakis2022misuses}) to
the precipitation time series from each of these locations to test if they
follow the
Generalized Extreme Value (GEV) distribution using the \textsl{evd} 
package \citep{evd}. Precipitation maximum series 
typically do follow the GEV distribution, so we expect to fail to reject the
null hypothesis.

\jy{Is the parametric method by Zeimberkakis?}
\mc{yes}

@@

\subsection{SP500}
\label{sec:sp500}

Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the SP500 index. We
observed how the method performs using increasingly smaller durations of
time.
We used the last 5 years of data: January 1st, 2019 to December 31st, 2023,
4 years of data: January 1st, 2020 to December 31st, 2023,
3 years of data: January 1st, 2021 to December 31st, 2023,
2 years of data: January 1st, 2022 to December 31st, 2023,
and 1 year of data: January 1st, 2023
to December 31st, 2023.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
The p-values for these tests are highlighted in 
Tables~\ref{table:SP5001}--\ref{table:SP5005}.

\input{tables/sp500_pval_1y}

\input{tables/sp500_pval_2y}

\input{tables/sp500_pval_3y}

\input{tables/sp500_pval_4y}

\input{tables/sp500_pval_5y}

While we expect there to be differences in the p-values for the different 
methods, in this example, there are many instances where the conclusions
at the .05 significance level for our method are different than those
for both
\citep{babu2004goodness}'s and \citet{zeimbekakis2022misuses}'s method.
For instance, for a duration of 4 years, our method fails to reject that
the series follows the $v = 5$ Student's $t$ distribution with p-value .064,
wheraeas \citep{babu2004goodness}'s method finds a p-value of .016, and 
\citet{zeimbekakis2022misuses}'s method finds a p-value of .01.

\subsection{SP100}
\label{sec:sp100}

Once again using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the SP100 index. We varied the
durations as we did with the SP500 index.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
The p-values for these tests are highlighted in 
Tables~\ref{table:SP1001}--\ref{table:SP1005}.
For a duration of 4 years, our method fails to reject that
the series follows the $v = 5$ Student's $t$ distribution with p-value .092,
wheraeas \citep{babu2004goodness}'s method finds a p-value of .031, and 
\citet{zeimbekakis2022misuses}'s method finds a p-value of .027.

\input{tables/sp100_pval_1y}

\input{tables/sp100_pval_2y}

\input{tables/sp100_pval_3y}

\input{tables/sp100_pval_4y}

\input{tables/sp100_pval_5y}
